Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2023-06-06 inst/manuscript/manuscript_microxanox.Rmd
References adjusted

linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output:
  rticles::elsevier_article:
    includes:
      in_header: headers.tex

# Citation settings
bibliography: bibliography.bib
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
---

@@

A typical simulation would look as shown in Figure \ref{fig:runsim_example}.

```{R runsim_example, eval = TRUE, echo = FALSE, fig.cap='Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.', fig.align="center", out.width = 350 }
fig <- "./figures/simflow.pdf"

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This function is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.

@@

An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}). The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria. Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types. The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.

@@

The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}). For example, with only the CB (cyanobacteria) present, there was no hysteresis. In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis. (These results are also given in the *Partial Reproduction* vignette.)

@@

Figure \ref{fig:uc3} shows a simulation with two species (strains) in each of the three functional groups. The ecosystem starts in the oxic state, though with relatively high abundance of each functional group. The strain of SB that is more tolerant to oxygen (SB_1) initially decreases in abundance, but then increases, and the other (SB_2) strain then becomes abundance and SB_1 declines. Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state. In contrast, if there are two identical strains with tolerance half way between those in Figure \ref{fig:uc3} the ecosystem remains in the oxic state.

@@

# TODO in RMD file
# TODO in LaTeX file
- [ ] fix citation style
- [ ] fix references to figures

>>>>

linenumbers: true
numbersections: true
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    keep_tex: true
    includes:
      in_header: headers.tex

# Citation settings
bibliography: bibliography.bib
csl: 'softwarex.csl'
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
---

@@

A typical simulation would look as shown in Figure \@ref(fig:runsimexample).

```{R runsimexample, eval = TRUE, echo = FALSE, fig.cap='Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.', fig.align="center", out.width = 350 }
fig <- "./figures/simflow.pdf"

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This function is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in Figure \@ref(fig:plot-dynamics).

@@

An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \@ref(fig:uc1)). The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria. Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types. The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.

@@

The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \@ref(fig:uc2)). For example, with only the CB (cyanobacteria) present, there was no hysteresis. In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis. (These results are also given in the *Partial Reproduction* vignette.)

@@

Figure \@ref(fig:uc3) shows a simulation with two species (strains) in each of the three functional groups. The ecosystem starts in the oxic state, though with relatively high abundance of each functional group. The strain of SB that is more tolerant to oxygen (SB_1) initially decreases in abundance, but then increases, and the other (SB_2) strain then becomes abundance and SB_1 declines. Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state. In contrast, if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3) the ecosystem remains in the oxic state.

@@

### 2023-06-06 inst/manuscript/manuscript_microxanox.tex
References adjusted

 \documentclass[preprint,12pt, a4paper]{elsarticle}

%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
% \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

\usepackage[unicode=true]{hyperref}

\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother

@@

@@

@@

  \begin{abstract}
  \(Microxanox\) is an R package to simulate a three functional group
  ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
  oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
  differential equations. Simulations can be run individually or over a
  parameter range, to find stable states. The model can be implemented
  with different numbers of species per functional group. The package is
  constructed in such a way that the results contain the input parameter
  used, so that a saved results can be loaded again and the simulation
  be repeated. Furthermore, the package framework and code should serve
  as a useful starting point for making simulation models of other types
  of ecosystems.
  \end{abstract}
    \begin{keyword}
    reproducibility \sep regime shift \sep stable state \sep 
    ordinary differential equations
  \end{keyword}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & v0.9.0 \\
C2 & Permanent links to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Owen.Petchey@ieu.uzh \\
\end{longtable}

@@

Mathematical models play a key role in the development of ecosystem
models and the understanding about how ecosystems work and respond to
environmental changes
\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
important for developing hypotheses to test in empirical studies. One
area of ecology in which models have been influential is the
understanding of their response to a gradual change in an environmental
driver \citep{Scheffer2001}. An environmental driver is an environmental
condition that affects an ecosystem. Still, it is assumed not to be
affected by the ecosystem, such as the rate of nutrient input into a
lake.

@@

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change \citet{Limberger2022}, and to base our work on the
work and model of \citet{Bush2017}. With this goal in mind, we developed
the microxanox package \citep{Krug2022}. The first stage of development
was to write code from scratch (as there was no available code to start
from) and to confirm that this new implementation would reproduce the
previously published results. The resulting reproduction is available as
one of the package vignettes:
\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
Partial reproduction of Bush et al}.

@@

The framework used when writing this package aimed to maximise
simplicity for the user and to make it straightforward to reproduce
results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
for an example of how this is used). As such, all the parameters needed
to run a simulation or find a stable state (i.e.~the final state of the
ecosystem) are contained in a single object (which can easily be created
using included functions). This parameter object is given to a function
that runs the simulations and returns the results. The returned results
object is identical to the parameter object but with an additional slot
named \texttt{results}, which contains the simulation results. Thus the
returned results object contains the simulation conditions (parameters)
as well as the results, and can be used to run the simulation again.
This promotes reproducibility and makes incremental changes of
individual parameters with a consecutive re-running of the simulations
straightforward.

@@

A typical simulation would look as shown in Figure
\ref{fig:runsim_example}.

@@

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
\end{figure}

@@

\hypertarget{finding-stable-states}{%
\subsection{Finding Stable States}\label{finding-stable-states}}

The general approach to finding the numerical value of a system's stable
state (if it exists) with a specific parameter set is to run the
simulation for a long time and record the final state. When one does
this across a range of environmental conditions, one discovers how the
steady state of the system responds to the environmental conditions. The
package contains functionality for finding steady states that correspond
to values of one environmental driver, namely the value of oxygen
diffusivity.

Two methods for finding steady states are implemented. The first runs a
separate simulation for each combination of starting conditions and
oxygen diffusivity (we term this the \emph{Replication method}). This is
the method used in the \citet{Bush2017} study. The second runs two
simulations, one with step-wise and slowly temporally \emph{increasing}
oxygen diffusivity, and the other with step-wise and slowly
\emph{decreasing} oxygen diffusivity. During this temporal environmental
change, the state of the system is recorded just before change to a new
oxygen diffusivity (we term this the \emph{Temporal method}). We
implemented two methods since there is no definitive best method, and in
order to check if results were sensitive to choice of method.

@@

From the results returned, summary measures about how the ecosystem
stable states respond to environmental change can be extracted. The
function \texttt{get\_stability\_measures()} returns quantities such as
the amount of environmental change required for the system to abruptly
change to a different state.

@@

The open source implementation and extension of the model used in
\citet{Bush2017} provides the means of reproducing the results published
while at the same time provides the means of doing unique, innovative,
and important investigations of how ecosystems respond to environmental
change, and how biodiversity may modulate this response.

@@

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from \citet{Limberger2022} and \citet{Petchey2022}.

@@

The study of \citet{Bush2017} includes simulations of the effect of
oxygen diffusivity (an environmental driver, in the sense that it
affects the ecosystem but is not affected by it) on the ecosystem state
(oxic or anoxic). The \emph{microxonax} package contains functionality
to make a specific temporal pattern of change in the oxygen diffusivity.
As well as allowing individual simulations during which oxygen
diffusivity varies, this functionality forms the basis of the temporal
method for finding stable states.

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

The package contains a function to extract summary features of ecosystem
responses to environmental change, such as the amount of hysteresis
displayed by the ecosystem. Hysteresis is a key feature of ecosystem
responses to environmental change, because it is related to how
difficult it can be to reserve the effects of environmental change
\citep{Scheffer2001}. The amount of hysteresis is measured as the extent
of the environmental condition (here oxygen diffusitivity) for which
there were two stable states. I.e. it is the extent of the environmental
conditions for which historical conditions play an important role in
determining the current system state (a definition of hysteresis).

Using the package to calculate the extent of hysteresis involves setting
ecosystem and simulation parameters, including parameters for the
finding of stable states across an environmental gradient, running the
stable state finding function, and analysing the results with the
function that calculates extent of hysteresis. The code for this is
provided in the \emph{User Guide}.

@@

\renewcommand\refname{References}
\bibliography{bibliography}
\bibliographystyle{elsarticle-num-names.bst}

>>>>

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

@@

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

@@

%\usepackage{lineno}
%\linenumbers
%\usepackage{setspace}
%\doublespacing
%\usepackage{fancyhdr}
%\fancyhead{}
%\fancyfoot{}
%\fancyhead[CO,CE]{Microxanox R package}
%\pagestyle{fancy}
%

@@

  \begin{abstract}
  \(Microxanox\) is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations. Simulations can be run individually or over a parameter range, to find steady states. The model can be implemented with different numbers of species per functional group. The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated. Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
  \end{abstract}
    \begin{keyword}
    reproducibility \sep regime shift \sep steady state \sep 
    ordinary differential equations
  \end{keyword}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & v0.9.1 \\
C2 & Permanent links to code/repository used for this code version & \url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/7148667}{10.5281/zenodo.7148667} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & \href{https://git-scm.com}{git} \\
C6 & Software code languages, tools, and services used & \href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments & \href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual & \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh \\
\end{longtable}

@@

Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also important for developing hypotheses to test in empirical studies. One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver \citep{Scheffer2001}. An environmental driver is an environmental condition that affects an ecosystem. Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.

@@

We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change \citet{Limberger2022}, and to base our work on the work and model of \citet{Bush2017}. With this goal in mind, we developed the \texttt{microxanox} package \citep{Krug2022}. The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results. The resulting reproduction is available as one of the package vignettes: \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette Partial reproduction of Bush et al}.

@@

The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2022} for an example of how this is used). As such, all the parameters needed to run a simulation or find a steady state (i.e.~the final state of the ecosystem) are contained in a single object (which can easily be created using included functions). This parameter object is given to a function that runs the simulations and returns the results. The returned results object is identical to the parameter object but with an additional slot named \texttt{results}, which contains the simulation results. Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again. This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.

@@

A typical simulation would look as shown in Figure \ref{fig:runsimexample}.

@@

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsimexample}
\end{figure}

@@

\hypertarget{finding-steady-states}{%
\subsection{Finding Steady States}\label{finding-steady-states}}

The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state. When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions. The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.

Two numerical strategies for finding steady states and their sensitivity to parameters are implemented. Two strategies are implemented in order to allow comparison of their results. The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the \emph{Replication method}). This is the method used in the \citet{Bush2017} study and was used to obtain the results in figures 3 and 4 of that article.

The second method runs two simulations, one with step-wise and slowly temporally \emph{increasing} oxygen diffusivity, and the other with step-wise and slowly \emph{decreasing} oxygen diffusivity. (Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated. Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the \emph{Temporal method}).

The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.

@@

From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted. The function \texttt{get\_stability\_measures()} returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.

@@

The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.

@@

Here we evidence the impact of the \emph{microxanox} package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} and the \emph{Partial Reproduction} vignettes. The third is taken from \citet{Limberger2022} and \citet{Petchey2022}.

@@

The study of \citet{Bush2017} includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic). The \texttt{microxonox} package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity. As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

The package contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem. Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change \citep{Scheffer2001}. The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states. I.e. it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).

Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis. The code for this is provided in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.

@@

\renewcommand\refname{References}
\bibliography{bibliography.bib}
